An Online Visual Loop Closure Detection Method for Indoor Robotic Navigation

### 2014-12-02 latex-spie/article.tex
almost finished

Visual loop closing techniques can be categorized into three main parts\cite{Williams09}: Image-to-image and image-to-map, map-to-map techniques. Loop closure estimations\cite{sariyanidi_spie13}  are cast by matching images. Most of image-to-image techniques that rely on salient image patches, use small, low-level features such as SURF\cite{springerlink:10.1007/11744023_32} extracted from the whole image \cite{sariyanidi_iv, HoNewmanCIRA2005, CumminsNewmanIJRR08}. One of the widely used methods in visual place recognition problem is the Bag of Words\cite{1238663} method. Also, there are some studies that uses additional depth information\cite{kerl2013-IROS} in loop closure problems.More information related to visual loop closing approaches can be found on the relevant survey \cite{Williams09}.

@@

The technique we implement relies on discrete Complex Zernike Moments (CZMs) in 2-dimensional space. They are extracted using a set of complex polynomials which form a complete orthogonal radial basis functions defined on the unit disc. These moments are used to represent shape information inside the image within the context of image processing. To make the image description more robust, the CZMs of the input image is calculated in local manner. In this study, the CZMs are extracted from both the greyscale and depth images separately and concatenated sequentially to create ultimate feature vector.

The locations are represented with the histograms of their images. Detecting loop closing events can be considered as a machine learning problem which number of classes expands continuously when a new unseen location comes. We utilized Nearest Neighbour (NN) algorithm to classify locations by which the distance metric is the regular $L_1$ a.k.a. Manhattan distance that is one of the simplest distance metric available. In other words, the image which distance between the input image is minimum closes the loop, if the distance is lower than the predefined threshold value.

@@

The method implemented in this paper has been evaluated on three datasets which lighting conditions are different from each other. The tests are done with either using gray and depth images separately or using both of them. After that, the difference matrices are calculated to examine the results. The NN algorithm is applied to classify the new locations using the $L_{1}$ distance metric. In addition to this, the last 100 frames are not included in the classification phase, because it the sensor is always moving and these frames correspond to the area behind the sensor.

@@

        %%% BRIGHT PR CURVE
        \begin{subfigure}[b]{0.48\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_bright.eps}    
        \caption{Bright}
        \label{subfig:pr_bright}
        \end{subfigure}
        ~ 
        %%% DIM PR CURVE
        \begin{subfigure}[b]{0.48\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_dim.eps}    
        \caption{Dim}
        \label{subfig:pr_dim}
        \end{subfigure}
        ~
        %%% DARK PR CURVE
        \begin{subfigure}[b]{0.48\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_dark.eps}    
        \caption{Dark}
        \label{subfig:pr_dark}
        \end{subfigure}
        ~ 
        %%% ALL PR CURVE
        \begin{subfigure}[b]{0.48\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_all.eps}    
        \caption{Comparison between the datasets.}
        \label{subfig:pr_all}
        \end{subfigure}

@@

The results are obtained by three datasets for which the trajectory is similar. Each dataset that is created in different lightning conditions: Bright, Dim and Dark. In the first loop, there are no loop closures to detect. To examine the performance of the method, precision-recall (PR) curves are plotted for each dataset. The PR curves are illustrated in Figures \ref{subfig:pr_bright}, \ref{subfig:pr_dim}, \ref{subfig:pr_dark} in which there are compared three different image sources that are only grayscale images, only depth images and both. The results show that using both grayscale and depth images gives the best performance in bright and dim datasets. In addition to this, the similarity matrices calculated from each dataset is shown in Figure \ref{fig:dist_matrices}. The black dots corresponds to the detected loop closures at $90\%$ precision. As it can be seen in the similarity matrices, the algorithm performs quite successful results to detect loop closure event.

%VANGELden
%the performance shown in Table/figure is  representative as the results are obtained with a %manually annotated ….
%our demo video provides qualitative loop closure results which may be considered to be more representative for our method’s performance

@@

        %%% BRIGHT DISTANCE MATRIX
        \begin{subfigure}[b]{0.27\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{figures/dist_bright.png}    
        \caption{Bright}
        \label{subfig:dist_bright}
        \end{subfigure}
        ~ 
        %%% DIM DISTANCE MATRIX
        \begin{subfigure}[b]{0.27\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{figures/dist_dim.png}    
        \caption{Dim}
        \label{subfig:dist_dim}
        \end{subfigure}
         ~ 
        %%% DARK DISTANCE MATRIX
        \begin{subfigure}[b]{0.27\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{figures/dist_dark.png}    
        \caption{Dark}
        \label{subfig:dist_dark}
        \end{subfigure}

>>>>

Visual loop closing techniques can be categorized into three main parts\cite{Williams09}: Image-to-image and image-to-map, map-to-map techniques. Loop closure estimations\cite{sariyanidi_spie13}  are cast by matching images. Most of image-to-image techniques that rely on salient image patches, use small, low-level features such as SURF\cite{springerlink:10.1007/11744023_32} extracted from the whole image \cite{sariyanidi_iv, HoNewmanCIRA2005, CumminsNewmanIJRR08}. One of the widely used methods in visual place recognition problem is the Bag of Words\cite{1238663} method. Also, there are some studies that uses additional depth information\cite{kerl2013-IROS} in loop closure problems. More information related to visual loop closing approaches can be found on the relevant survey \cite{Williams09}.

@@

The technique we implement relies on discrete complex Zernike moments (CZMs) in 2-dimensional space. They are extracted using a set of complex polynomials which form a complete orthogonal radial basis functions defined on the unit disc. These moments are used to represent shape information inside the image within the context of image processing. To make the image description more robust, the CZMs of the input image is calculated in local manner. In this study, the CZMs are extracted from both the greyscale and depth images separately and concatenated sequentially to create ultimate feature vector.

The locations are represented with the histograms of their images. Detecting loop closing events can be considered as a machine learning problem which number of classes expands continuously when a new unseen location comes. We utilized nearest neighbour algorithm to classify locations by which the distance metric is the regular $L_1$ a.k.a. Manhattan distance that is one of the simplest distance metric available. In other words, the image which distance between the input image is minimum closes the loop, if the distance is lower than the predefined threshold value.

@@

The method implemented in this paper has been evaluated on three datasets which lighting conditions are different from each other. The tests are done with either using grayscale and depth images separately or using both of them. After that, the difference matrices are calculated to examine the results. The NN algorithm is applied to classify the new locations using the $L_{1}$ distance metric. In addition to this, the last 100 frames are not included in the classification phase, because it the sensor is always moving and these frames correspond to the area behind the sensor.

@@

        %%% BRIGHT PR CURVE
        \begin{subfigure}[b]{0.43\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_bright.eps}    
        \caption{Bright}
        \label{subfig:pr_bright}
        \end{subfigure}
        ~ 
        %%% DIM PR CURVE
        \begin{subfigure}[b]{0.43\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_dim.eps}    
        \caption{Dim}
        \label{subfig:pr_dim}
        \end{subfigure}
        ~
        %%% DARK PR CURVE
        \begin{subfigure}[b]{0.43\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_dark.eps}    
        \caption{Dark}
        \label{subfig:pr_dark}
        \end{subfigure}
        ~ 
        %%% ALL PR CURVE
        \begin{subfigure}[b]{0.43\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, width=\textwidth]{figures/pr_all.eps}    
        \caption{Comparison between the datasets.}
        \label{subfig:pr_all}
        \end{subfigure}

@@

The results are obtained from three datasets for which the trajectory of the testing platform was the same. Each dataset is gathered in different lighting conditions: bright, dim and dark.
To examine the performance of the method, precision-recall (PR) curves are plotted for each dataset. The figures \ref{subfig:pr_bright}, \ref{subfig:pr_dim}, \ref{subfig:pr_dark} consist of PR curves belonging to test runs in which only grayscale images are present, only depth-images are present or both grayscale and depth images are present. In Figure \ref{subfig:pr_all}, using the test run done with both grayscale and depth images, a comparison plot is shown, consisting of three PR curves belonging to bright, dim and dark datasets. The results show that using both grayscale and depth images gives the best performance in bright and dim datasets. In addition to this, the similarity matrices calculated from each dataset is shown in Figure\ref{fig:dist_matrices}. The black dots corresponds to the detected loop closures at $90\%$ precision. As it can be seen in the similarity matrices, the algorithm accomplishes to detect loop closure events.

@@

        %%% BRIGHT DISTANCE MATRIX
        \begin{subfigure}[b]{0.30\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{figures/dist_bright.png}    
        \caption{Bright}
        \label{subfig:dist_bright}
        \end{subfigure}
        ~ 
        %%% DIM DISTANCE MATRIX
        \begin{subfigure}[b]{0.30\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{figures/dist_dim.png}    
        \caption{Dim}
        \label{subfig:dist_dim}
        \end{subfigure}
         ~ 
        %%% DARK DISTANCE MATRIX
        \begin{subfigure}[b]{0.30\textwidth}
        \centering
         % trim parameters by order: left, bottom, right, top
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{figures/dist_dark.png}    
        \caption{Dark}
        \label{subfig:dist_dark}
        \end{subfigure}
